Baseline centering problem

### 2018-12-06 paper/tangotd.tex

@@

A few choices are:
\begin{itemize}
\item GTD is a gradient descent of
$\transp{(A\theta-b)}(A\theta-b)$ (Sutton 2009).
This corresponds to $B=\transp{A}$. Then
$BA=\transp{A}A$ is
obviously stable. This helps with $A$ being non-symmetric. TODO: check if
GTD uses the same sample for $\tilde{A}$ and $A$. If so, it will
converge to a different stable point. TODO: Can be applied TANGO-style
(by formally inverting $\Id$ in TANGO), advantage=fewer variance due to
two-sampling (Is this equivalent to eligibility traces without the
variance??). (TANGO uses two-sampling but on a sum of past terms, so
that a past term gets TANGOified whenever we happen to reach a state that
activates the same features in the future, which is probably fine). Can
also be TANGOified with $\gamma\to 0$ (two-timescale) which now requires
only single sampling (no replay buffers).

@@

[Combined with all these problems, is the problem that a Bellman solution on the
support of $\rho$ may be very, very far from the global Bellman solution,
even on the support of $\rho$ (take random walk on a circle, with $\rho$
being everything but one point: then any linear function solves Bellman
on the support of $\rho$).]

>>>>

\section{Recovering known convergence results for TD, and a few ideas}

@@

A few choices are:
\begin{itemize}
\item GTD is a gradient descent of
$\transp{(A\theta-b)}(A\theta-b)$ (Sutton 2009).
This corresponds to $B=\transp{A}$. Then
$BA=\transp{A}A$ is
obviously stable. This helps with $A$ being non-symmetric. TODO: check if
GTD uses the same sample for $\transp{A}$ and $A$. If so, it will
converge to a different stable point. TODO: Can be applied TANGO-style
(by formally inverting $\Id$ in TANGO), advantage=fewer variance due to
two-sampling (Is this equivalent to eligibility traces without the
variance??). (TANGO uses two-sampling but on a sum of past terms, so
that a past term gets TANGOified whenever we happen to reach a state that
activates the same features in the future, which is probably fine). Can
also be TANGOified with $\gamma\to 0$ (two-timescale) which now requires
only single sampling (no replay buffers).

@@

\section{A function that almost solves the Bellman equation can be very
far from the true Bellman function}

Combined with all these problems, is the problem that a Bellman solution on the
support of $\rho$ may be very, very far from the global Bellman solution,
even on the support of $\rho$ (take random walk on a circle, with $\rho$
being everything but one point: then any linear function solves Bellman
on the support of $\rho$).

TODO seems to be much worse off-policy.

\section{Storing a summary of possible transitions?}

The derivative of the average reward with respect to the policy
parameters $\psi$ is equal to
\begin{align}
\partial_\psi \left(\E_\mu r\right)&=
\E_{s,a,s'} \partial_\psi \log \pi(a|s)\left(
r+V(s')
\right)
\\&=\E_{s,a}\partial_\psi \log \pi(a|s)\E_{s'|s,a}\sum_{k\geq 0}P^k R(s')
\\&=\E_{s,a,s'}\partial_\psi \log \pi(a|s)(\Delta^{-1}R)(s')
\end{align}
by expanding $V(s')$ as a sum of expected future rewards (and attributing
the reward of a transition $s\to s'$ to $s'$, which makes sense if $s'$
is a sufficient statistic).
(Here I'm neglecting constant functions again.)

Thus if we denote
\begin{equation}
g(s')\deq \E_{s,a,s'}\partial_\psi \log \pi(a|s)\1_{s'}
\end{equation}
we have
\begin{equation}
\partial_\psi \left(\E_\mu r\right)=\transp{g}\Delta^{-1}R
\end{equation}
Let us define the row vector
\begin{equation}
G\deq \transp{g}\Delta^{-1},\qquad \partial_\psi \left(\E_\mu
r\right)=GR
\end{equation}
then one has
\begin{equation}
G(s)=\partial_\psi \mu(s)
\end{equation}
namely, $G$ encodes the derivative of the invariant distribution with
respect to the policy parameters. Thus, if we find a state $s$ with high
reward $R(s)$, $G$ tells us how we can visit $s$ more often. This
explains that the derivative of the average reward is $GR$.

Eligibility traces are a random, unbiased estimator of $G$: if $s_t$ is a
trajectory from the current policy then
\begin{equation}
\E(\sum_{k\geq 1}\partial_\psi \pi(a_{t-k}|s_{t-k})\1_{s_t})=G
\end{equation}
Moreover $G$ satisfies a backward Bellman equation on-policy,
\begin{equation}
G=\E_{s,a,s'} \partial_\psi \pi(a|s)\1_{s'}+GP
\end{equation}
thanks to $\Delta^{-1}=\sum P^k=\Id+\Delta^{-1}P$. Eligibility traces are
updated this way at each step, via a sampling of $P$.

$G$ cannot be stored but one can store an approximate model. For instance
one could approximate $G$ as a sum of parameters times functions:
\begin{equation}
G\approx \sum_i \psi_i \phi_{\theta_i}(s)
\end{equation}
where each $\psi_i$ is a policy parameter, and each $\phi_{\theta_i}$ is
a function of the state parameterized by some model. Then the reverse
Bellman equation above allows us to learn $\psi_i$ and $\theta_i$ via 
a TD-like algorithm.

TODO: not the right representation, plus works only on-policy. Maybe learn a function $a(s,s')$ that
returns the first action to be taken if we want to reach $s'$ from $s$?
